---
title: "Craft-Cans Case Study"
author: "Arturo Casillas & Kevin Dickens"
date: "October 6, 2017"
output: 
   #md_document:
   #  variant: markdown_github
   knitrBootstrap::bootstrap_document:
        theme: "spacelab"
        highlight: Visual Studio
        theme.chooser: TRUE
        highlight.chooser: TRUE
        #menu: FALSE
        custom.header: NULL
       # knit_bootstrap_md:(input="CaseStudy1BootstrapDoc.Rmd", output="CaseStudy1BootstrapDoc.md")
   # html_document:
   #  #template: quarterly_report.html
   #  toc: true
   #  toc_float: true
   #  code_folding: hide
   #  #df_print: kable
   #  #df_print: paged
   #  theme: "readable"
   #  keep_md: true
---

#Intro
##Intro

-Intro goes here


```{r echo=FALSE, include = FALSE}
#bootstrap.show.code=FALSE, bootstrap.show.message=FALSE, bootstrap.show.warning=FALSE
## Load Packages ##
library(ggplot2)
library(fiftystater)
library(plyr)
library(datasets)
library(maps)
library(mapproj)


library(knitr)
opts_chunk$set(bootstrap.show.code=FALSE, bootstrap.thumbnail=FALSE, bootstrap.show.message=FALSE)
library(ggplot2)
theme_set(new = theme_minimal())

#use color brewer as default discrete colors
scale_colour_discrete <- function(...) scale_color_brewer(palette="Set1", ...)
scale_fill_discrete <- function(...) scale_fill_brewer(palette="Set1", ...)

library(plyr)
## Load Data ##
#getwd()
rawbeers <- read.csv("C:/Users/acasi/Downloads/Beers.csv")
rawbreweries <- read.csv("C:/Users/acasi/Downloads/Breweries.csv")
library(xtable)
options(xtable.floating = FALSE)
options(xtable.timestamp = "")
library(pander)


##### State Name Function #####

StateName<-function(test){
  
  
  
  #Checks to factors, converts them to characters, and then trims any white space before or after the characters
  
  i <- sapply(test, is.factor)
  
  test[i] <- lapply(test[i], as.character)
  
  test$State<-sapply(test$State, trimws, which="both")
  
  
  
  #Replaces each state code with the corresponding state name.  Note that the df must have a column named State that contains the state codes for this code to work.
  
  test$State[test$State =="AL"]<-"alabama"
  
  test$State[test$State =="AK"]<-"alaska"
  
  test$State[test$State =="AZ"]<-"arizona"
  
  test$State[test$State =="AR"]<-"arkansas"
  
  test$State[test$State =="CA"]<-"california"
  
  test$State[test$State =="CO"]<-"colorado"
  
  test$State[test$State =="CT"]<-"connecticut"
  
  test$State[test$State =="DC"]<-"district of columbia"
  
  test$State[test$State =="DE"]<-"deleware"
  
  test$State[test$State =="FL"]<-"florida"
  
  test$State[test$State =="GA"]<-"georgia"
  
  test$State[test$State =="HI"]<-"hawaii"
  
  test$State[test$State =="ID"]<-"idaho"
  
  test$State[test$State =="IL"]<-"illinois"
  
  test$State[test$State =="IN"]<-"indiana"
  
  test$State[test$State =="IA"]<-"iowa"
  
  test$State[test$State =="KS"]<-"kansas"
  
  test$State[test$State =="KY"]<-"kentucky"
  
  test$State[test$State =="LA"]<-"louisiana"
  
  test$State[test$State =="ME"]<-"maine"
  
  test$State[test$State =="MD"]<-"maryland"
  
  test$State[test$State =="MA"]<-"massachusetts"
  
  test$State[test$State =="MI"]<-"michigan"
  
  test$State[test$State =="MN"]<-"minnesota"
  
  test$State[test$State =="MS"]<-"mississippi"
  
  test$State[test$State =="MO"]<-"missouri"
  
  test$State[test$State =="MT"]<-"montana"
  
  test$State[test$State =="NE"]<-"nebraska"
  
  test$State[test$State =="NV"]<-"nevada"
  
  test$State[test$State =="NH"]<-"new hampshire"
  
  test$State[test$State =="NJ"]<-"new jersey"
  
  test$State[test$State =="NM"]<-"new mexico"
  
  test$State[test$State =="NY"]<-"new york"
  
  test$State[test$State =="NC"]<-"north carolina"
  
  test$State[test$State =="ND"]<-"north dakota"
  
  test$State[test$State =="OH"]<-"ohio"
  
  test$State[test$State =="OK"]<-"oklahoma"
  
  test$State[test$State =="OR"]<-"oregon"
  
  test$State[test$State =="PA"]<-"pennsylvania"
  
  test$State[test$State =="RI"]<-"rhode island"
  
  test$State[test$State =="SC"]<-"south carolina"
  
  test$State[test$State =="SD"]<-"south dakota"
  
  test$State[test$State =="TN"]<-"tennessee"
  
  test$State[test$State =="TX"]<-"texas"
  
  test$State[test$State =="UT"]<-"utah"
  
  test$State[test$State =="VT"]<-"vermont"
  
  test$State[test$State =="VA"]<-"virginia"
  
  test$State[test$State =="WA"]<-"washington"
  
  test$State[test$State =="WV"]<-"west virginia"
  
  test$State[test$State =="WI"]<-"wisconsin"
  
  test$State[test$State =="WY"]<-"wyoming"
  
  
  
  #returns the new dataframe with the state names.

  return(test)
  
}
```

#Data
##About the Data

###Source

The tables contain a list of 2410 US craft beers and 510 US breweries. The beer data corresponds to craft beers available in cans and lists the beer ID number, the size of the can in ounces, style of beer, percent alcohol per volume (ABV), and international bitterness units (IBU) as well as the beer name and brewery ID.The Breweries data lists breweries by location of state and city along with a unique ID. This data was traced to CraftCans.com and further traced to the Brewers Association (BA). A more expansive data set is available to Brewers Association members. 

To prepare the data for analysis, variable names are altered for clarity and to minimize merging issues. The beers and breweries are linked by a unique numeric ID (Brew_ID), so no assumptions or algorithm was needed to combine the tables. 

-View the original data. Expand below to see details.

```{r view, bootstrap.show.code=FALSE, bootstrap.show.message=FALSE, bootstrap.show.warning=FALSE, bootstrap.show.output=FALSE}

str(rawbeers)

str(rawbreweries)

```

-Prepare for merging. Expand below to see details.

```{r prep, bootstrap.show.code=FALSE, bootstrap.show.message=FALSE, bootstrap.show.warning=FALSE, bootstrap.show.output=FALSE}
## Prepare for merging ##

names(rawbeers)[5] = "Brew_ID"  #Rename 'Brewery_id' to ease merging

names(rawbeers)[1] = "Beer_Name" #Rename 'Name' to 'Beer_Name' to avoid confusion with brewery name or state

names(rawbreweries)[2] = "Brewery_Name"  #Rename 'Name' to 'Brewery_Name' avoid confusion and ease merging

```

-Merged to create a single table, details below.

```{r merge, bootstrap.show.code=FALSE, bootstrap.show.message=FALSE, bootstrap.show.warning=FALSE, bootstrap.show.output=FALSE}
## Merge Data ##

AllBeer <- merge(rawbeers, rawbreweries, by="Brew_ID")

str(AllBeer)

levels(AllBeer$State)=trimws(levels(AllBeer$State), which = c("left"))  #Trim the leadig spaces from state abbreviation
```

-Expand below to see the first 6 rows and perform a spot check for integrity (raw)
```{r head, bootstrap.show.code=FALSE, bootstrap.show.message=FALSE, bootstrap.show.warning=FALSE, bootstrap.show.output=FALSE}
#kable(head(AllBeer, 6), row.names = FALSE)
head(AllBeer, 6)
```

-Also expand below to see the last 6 rows and perform a spot check for integrity (pander)
```{r tail, bootstrap.show.code=FALSE, bootstrap.show.message=FALSE, bootstrap.show.warning=FALSE, bootstrap.show.output=FALSE}
panderOptions('table.split.table', 100)
panderOptions('table.continues', '...')
pander(tail(AllBeer, 6), row.names = FALSE)
```

###Data Integrity

The data were examined for missing values and nonsense entries through formal methods in addtion to spot checks. In summary, Only Style, ABV and IBU have missing values. Almost 50% of IBU values are missing, which will most certainly affect our conclusions. Only 3% of ABV values and 5 Style entries are missing. 

-Confirm numeric missing values
```{r num.missing, bootstrap.show.code=FALSE, bootstrap.show.message=FALSE, bootstrap.show.warning=FALSE, bootstrap.show.output=FALSE}
## Prepare for merging ##

panderOptions('table.split.table', 100)
panderOptions('table.continues', '')
pander(apply(apply(AllBeer, 2, is.na), 2, sum))
```

-Confirm character missing values
```{r char.missing, bootstrap.show.code=FALSE, bootstrap.show.message=FALSE, bootstrap.show.warning=FALSE, bootstrap.show.output=FALSE}
## Double Check ##
## Look at blank Strings ##

panderOptions('table.split.table', 100)
panderOptions('table.continues', '')
pander(apply(AllBeer, 2, function(y) sum(y == "")))
```

###Add Region & Division
R contains additional region and division data per state in the package 'datasets'. These data are from a 1977 report from the chamber of commerce. The state classifications for region and division have not changed. Prior to June 1984 however, the Midwest Region was designated as the North Central Region. These data were added to expand analysis options and segmenting the states and cities geographically. 

```{r reg1, bootstrap.show.code=FALSE, bootstrap.show.message=FALSE, bootstrap.show.warning=FALSE, bootstrap.show.output=FALSE}
## Add region/division ##

state.geo=data.frame(state.abb, state.region, state.division)

#rename for ease of merge
#length(levels(state.geo$state.abb))
levels(state.geo$state.abb) = c(levels(state.geo$state.abb),"DC")
names(state.geo)[1]='State'
state.geo[51, ] = c('DC','South', 'South Atlantic')

#align levels for merge
levels(state.geo$State)=levels(AllBeer$State)


#Final Data#
levels(AllBeer$State)=trimws(levels(AllBeer$State), which = c("left"))
AllBeerReg<-merge(x=AllBeer, y=state.geo, by.x="State", all.x = TRUE)
str(AllBeerReg)

apply(AllBeerReg, 2, function(y) sum(y == ""))
apply(apply(AllBeerReg, 2, is.na), 2, sum)

```

#Analysis

###Syles

-The ten most popular canned beer styles
```{r pop1.Brew.H, bootstrap.show.code=FALSE, bootstrap.show.message=FALSE, bootstrap.show.warning=FALSE, bootstrap.show.output=FALSE}

#10 most popular beer styles
##use table to get frequencies and then sort
pop1=sort(table(AllBeerReg$Style))[100:90]
pop1

```

-The total number of breweries per state (horizontal)

```{r state.Brew.H, bootstrap.show.code=FALSE, bootstrap.show.message=FALSE, bootstrap.show.warning=FALSE, bootstrap.show.output=FALSE}
#Horizontal Table#

StBrews=sapply(tapply(AllBeerReg$Brew_ID, AllBeerReg$State, unique), length)
pander(StBrews)
names(rawbreweries)[2] = "Brewery_Name"  

```

- The total number of breweries per state (vertical)

```{r state.Brew.V, bootstrap.show.code=FALSE, bootstrap.show.message=FALSE, bootstrap.show.warning=FALSE, bootstrap.show.output=FALSE}

#Initializes a new object with State, Brewery Name, and City

newbrewery<-AllBeerReg[,c("Brewery_Name", "City", "State")]
str(newbrewery)

newbrewery2<-aggregate(newbrewery$Brewery_Name, list(newbrewery$State), function(x) length(unique(x)))   #Returns a count of the unique breweries by state

names(newbrewery2) = c("State", "Freq")

```

-Map of total number of breweries per state

```{r state.Brew.map, bootstrap.show.code=FALSE, bootstrap.show.message=FALSE, bootstrap.show.warning=FALSE, bootstrap.show.output=FALSE}
#Runs the StateName function to replace abbreviations with state names and then calls tolower to make certain they are lower case

newbrewery2<-StateName(newbrewery2)

newbrewery2$State<-tolower(newbrewery2$State)

str(newbrewery2)


# Start of the call to vreate the median IBU by state map

brew_map<- ggplot(newbrewery2, aes(map_id=State))+    #sets the data (newbrewery2) and the primary key (State variable) to link map and data
  
  geom_map(aes(fill=Freq), map=fifty_states)+         #sets the fill value (Freq) that will determine color and the geographic map data
  
  expand_limits(x=fifty_states$long, y=fifty_states$lat)+ #Sets the latitude and longitudinal extents
  
  coord_map() +                        #Sets the base geographic projection (mercator in this case)
  
  scale_x_continuous(breaks=NULL)+
  
  scale_y_continuous(breaks=NULL)+
  
  labs(x = "", y = "") +
  
  ggtitle("Number of Breweries By State") + # Sets the title of the map
  
  scale_fill_gradient(low = "seagreen1", high = "seagreen4", space = "Lab",na.value = "gray70", guide=guide_colourbar(title.position="top", barwidth=10, title="Number of Breweries",  title.hjust=0.5))+     #contols legend elements such as color gradiant, colors for NA values, and the size of the legend bar
  
  theme(plot.title = element_text(lineheight=.8, face="bold"),legend.position=c(.87, .25),legend.direction="horizontal",panel.background=element_blank(), panel.border=element_rect(colour="Grey50", fill=NA, size=2))+   #Theme elements such as the border around the map plot, the position of map components like the legend
  
  borders("state")+     #Adds colored state borders
  
  fifty_states_inset_boxes()     #Creates the insert boxes around Alaska and Hawaii so people don't mistake them for Mexican states or other Central American countries.



#Plots the map for total number of breweries by state

brew_map
  

```

 - Map of total number of breweries per state
 
### ABV and IBU Summary

Alcohol per volume and international bitterness units.

ABV does not have much variation. 'Lee Hill Series Vol. 5 Belgian Style Quadrupel Ale' out of Colorado is the most alcoholic beer. Six out of the ten most alcoholic beers are 'Imperial' beers while two out of the five most alcoholic beers come from 'Upslope Brewing Company' out of Boulder Colorado. Similarly, four of the ten most alcoholic beers come from California.

-Summary Statistics for ABV below
```{r summ.ABV, bootstrap.show.code=FALSE, bootstrap.show.message=FALSE, bootstrap.show.warning=FALSE, bootstrap.show.output=FALSE}
summary(AllBeerReg$ABV)
```

-List of 10 most alcoholic beers
```{r shigh.ABV, bootstrap.show.code=FALSE, bootstrap.show.message=FALSE, bootstrap.show.warning=FALSE, bootstrap.show.output=FALSE}
o1<-AllBeerReg[order(AllBeerReg$ABV, decreasing = TRUE), c("State", "Beer_Name", "ABV","IBU", "Brewery_Name", "City")]
kable(o1[1:10,])
#o1[1:10,]
```


The IBU distribution is very right skewed. The most bitter bear 'Bitter Bitch Imperial IPA' out of Oregon is the most bitter beer. Nine out of the 10 most bitter beers are IPAs and eight our of the 10 most bitter beers are Imperial IPAs. Also, there is no overlap between the 10 most bitter and 10 most alcoholic canned beers. However, all ten of the most bitter beers are above the median ABV of 0.056% and only one is below the mean ABV of 0.05977%. Similarly, all ten of the most alcoolic beers core above the mean and median IBU. 

-Summary statistics for IBU
```{r summ.IBU, bootstrap.show.code=FALSE, bootstrap.show.message=FALSE, bootstrap.show.warning=FALSE, bootstrap.show.output=FALSE}
summary(AllBeerReg$IBU)
```

-The 10 most bitter beers
```{r shigh.IBU, bootstrap.show.code=FALSE, bootstrap.show.message=FALSE, bootstrap.show.warning=FALSE, bootstrap.show.output=FALSE}
o2<-AllBeerReg[order(AllBeerReg$IBU, decreasing = TRUE),c("State", "Beer_Name", "ABV","IBU", "Brewery_Name", "City")]
kable(o2[1:10,])
#o2[1:10,]
```


ABV and IBU appear to be related, which is reasonable. The association is strong but not very powerful, in the sense that a large increase of IBU values are associated with a small increase in ABV despite being very likely to be associated with higher ABV.

-Barplot of median ABV and IBU per state
```{r med.ABV.IBU, bootstrap.show.code=FALSE, bootstrap.show.message=FALSE, bootstrap.show.warning=FALSE, bootstrap.show.output=FALSE, fig.width=12}

median_beer<-aggregate(AllBeerReg[,c("ABV","IBU")], list(AllBeerReg$State), median, na.rm=TRUE)    #Creates a new object which stores the median IBU and ABV per state
names(median_beer)[1]="State"


#barplot of median ABV and IBU per state in one graphic

par(mfrow=(c(2,1)), las=2, mar=c(2,2,1,1))

barplot(median_beer$ABV, names.arg = median_beer$State, cex.names = 0.7, cex.axis = 0.9, main="Median ABV by State", ylab="Alcohol By Volume (ABV)", las=2, col=c("lightgoldenrod2")) #ABV Bar Plot

barplot(median_beer$IBU, names.arg = median_beer$State, cex.names = 0.7, cex.axis = 0.9, main="Median IBU by State", ylab="International Bitterness Units (IBU)", las=2, col=c("lightblue3")) #IBU Bar Plot

```


-Map of median ABV per state
```{r map.ABV, bootstrap.show.code=FALSE, bootstrap.show.message=FALSE, bootstrap.show.warning=FALSE, bootstrap.show.output=FALSE}
median_beer2<-median_beer

median_beer2<-StateName(median_beer2)


# Start of the call to create the median ABV by state map

abv_map<- ggplot(median_beer2, aes(map_id=State))+      #sets the data and the primary key to link map and data
  
  geom_map(aes(fill=ABV), map=fifty_states)+                #sets the fill value that will determine color and the geographic map data
  
  expand_limits(x=fifty_states$long, y=fifty_states$lat)+   #Sets the latitude and longitudinal extents
  
  coord_map()+                                              #Sets the base geographic projection (mercator in this case)
  
  scale_x_continuous(breaks=NULL)+                          
  
  scale_y_continuous(breaks=NULL)+                          
  
  labs(x = "", y = "") +
  
  ggtitle("Median Alcohol By Volume (ABV) By State") +      # Sets the title of the map
  
  scale_fill_gradient(low = "lightgoldenrod1", high = "lightgoldenrod4", space = "Lab",na.value = "grey70", guide = guide_colourbar(title.position="top", barwidth=10, title.hjust=0.5))+ #contols legend elements such as color gradiant, colors for NA values, and the size of the legend bar
  
  theme(plot.title = element_text(lineheight=.8, face="bold"),legend.position=c(.875, .27),legend.text=element_text(size=8),legend.direction="horizontal", legend.margin=margin(t = 0, unit='cm'), panel.background=element_blank(), panel.border=element_rect(colour="Grey50", fill=NA, size=2))+   #Theme elements such as the border around the map plot, the position of map components like the legend
  
  borders("state")+                                         #Adds colored state borders
  
  fifty_states_inset_boxes()                                #Creates the insert boxes around Alaska and Hawaii so people don't mistake them for Mexican states or other Central American countries.



#Plots the ABV map

abv_map
```

-Map of median IBU per state
```{r map.IBU, bootstrap.show.code=FALSE, bootstrap.show.message=FALSE, bootstrap.show.warning=FALSE, bootstrap.show.output=FALSE}


# Start of the call to create the median IBU by state map

ibu_map<- ggplot(median_beer2, aes(map_id=State))+           #sets the data and the primary key to link map and data
  
  geom_map(aes(fill=IBU), map=fifty_states)+              #sets the fill value that will determine color and the geographic map data
  
  expand_limits(x=fifty_states$long, y=fifty_states$lat)+ #Sets the latitude and longitudinal extents
  
  coord_map()+                                            #Sets the base geographic projection (mercator in this case)
  
  scale_x_continuous(breaks=NULL)+
  
  scale_y_continuous(breaks=NULL)+
  
  labs(x = "", y = "") +
  
  ggtitle("Median International Bitter Units (IBU) By State") + # Sets the title of the map
  
  scale_fill_gradient(low = "lightcyan2", high = "lightblue4", space = "Lab",na.value = "gray70", guide = guide_colourbar(title.position="top", barwidth=10, title.hjust=0.5))+ #contols legend elements such as color gradiant, colors for NA values, and the size of the legend bar
  
  theme(plot.title = element_text(lineheight=.8, face="bold"),legend.position=c(.87, .25),legend.direction="horizontal",panel.background=element_blank(), panel.border=element_rect(colour="Grey50", fill=NA, size=2))+    #Theme elements such as the border around the map plot, the position of map components like the legend
  
  borders("state")+                #Adds colored state borders
  
  fifty_states_inset_boxes()        #Creates the insert boxes around Alaska and Hawaii so people don't mistake them for Mexican states or other Central American countries.



#Plots the IBU map

ibu_map

```


-Scatter plot of ABV and IBU
```{r ABV.v.IBU, bootstrap.show.code=FALSE, bootstrap.show.message=FALSE, bootstrap.show.warning=FALSE, bootstrap.show.output=FALSE}
ggplot(dat=AllBeerReg, aes(x=IBU, y=ABV)) + 
  geom_point(shape=16) + 
  geom_smooth(method=lm) + theme_minimal() 


```

-ABV to IBU regression equation
```{r ABV.v.IBU.eq, bootstrap.show.code=FALSE, bootstrap.show.message=FALSE, bootstrap.show.warning=FALSE, bootstrap.show.output=FALSE}
   L=lm(ABV~IBU, data=AllBeerReg)
  # L$coeff
  bquote(ABU[i] == IBU[i]%*% .(signif(L$coeff[2], 3)) + .(signif(L$coeff[1], 3)))
```

##Summary and Conclusions

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
