---
title: "Untitled"
author: "Arturo Casillas"
date: "October 6, 2017"
output: 
  html_document:
    keep_md: true
---

## R Markdown Dummy

- The tables contain a list of 2410 US craft beers and 510 US breweries respectively. The beers and breweries are linked together with an "id". This data was criginally from CraftCans.com. 

```{r cars}
## Load Data ##
getwd()
setwd("C:/Users/acasi/Downloads")
BeersData <- read.csv("Beers.csv")
BreweriesData <- read.csv("Breweries.csv")

str(BeersData)
str(BreweriesData)

## Prepare for merging ##
names(BeersData)
names(BeersData)[5] = "Brew_ID"
names(BeersData)[1] = "Beer.name"
names(BreweriesData)[2] = "Brewery.name"


## Merge Data ##
AllBeer <- merge(BeersData, BreweriesData, by="Brew_ID")
str(AllBeer)
```

## Headers & Summary

```{r head, echo=FALSE}
head(AllBeer, 6)
```

```{r tail, echo=FALSE}
tail(AllBeer, 6)
```

## Including Plots

Simple Plot

```{r plot1, echo=FALSE}
plot(ABV~IBU, data=AllBeer)
abline(lm(ABV~IBU, data=AllBeer))
```


Simple Plot

```{r plot2, echo=FALSE}
library(ggplot2)
ggplot(dat=na.omit(AllBeer), aes(x=IBU, y=ABV)) + 
  geom_point(shape=16) + 
  geom_smooth(method=lm) + theme_minimal()

#abline(lm(ABV~IBU, data=AllBeer))
```


Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
