---
title: "Craft-Cans Case Study"
author: "Arturo Casillas & Kevin Dickens"
date: 
output: 
  beamer_presentation:
    theme: "AnnArbor"
    colortheme: "dolphin"
    fonttheme: "structurebold"
#keep_tex: true

---

# R Markdown Dummy

First

# About the Data
## Source

- The tables contain a list of 2410 US craft beers and 510 US breweries. 
- The beer data corresponds to beers available in cans
- This data was traced from CraftCans.com. 

# About the Data
## View the data

```{r load}
## Load Data ##
#getwd()
setwd("C:/Users/acasi/Downloads")
BeersData <- read.csv("Beers.csv")
BreweriesData <- read.csv("Breweries.csv")

str(BeersData)
str(BreweriesData)
```

# Munging
## Basic alterations
- Column names are altered for clarity and to minimize merging issues
- The beers and breweries are linked by a numeric ID, which is used to merge the two tables
  - 'Brew_ID'

```{r munge}
## Prepare for merging ##
names(BeersData)
names(BeersData)[5] = "Brew_ID"
names(BeersData)[1] = "Beer.name"
names(BreweriesData)[2] = "Brewery.name"


## Merge Data ##
AllBeer <- merge(BeersData, BreweriesData, by="Brew_ID")
str(AllBeer)

```

# Munging
## View the final data

-first 6 rows
```{r head, fig.align='center', error=FALSE}
library(knitr)
#library(kableExtra)
kable(head(AllBeer, 6))
#%>%
#kable_styling(latex_options = c("striped", "scale_down"))
```

# Munging
### View the final data

last 6 rows

```{r tail, echo=TRUE}
library(knitr)
kable(tail(AllBeer, 6))
```


# Data Integrity
## Missing Vlaues
- Check the missing values number and indicators
```{r missing, echo=TRUE}
## Check for missing values ##
apply(apply(AllBeer, 2, is.na), 2, sum)

t=sapply(AllBeer, function(y) sum(length(which(is.na(y)))))
summary(t)
```

# Data Integrity
## Notes and Observations
- Only ABV and IBU have missing values
  - Almost 50% of IBU values are missing
  -This certainly affect any analysis that hinges on IBU values
- Spot checks on the data suggest that character variables use blanks for missing values
  - This needs to be verified

# Data Integrity
## Double Check Missing Vlaues
- Check for blanks instead of NA's
```{r missing2, echo=TRUE}
## Double Check ##
## Look at balnk Strings ##
apply(AllBeer, 2, function(y) sum(y == ""))
```
- 'Stye', which represents style of beer, has 5 additional missing values

# External Data

- From R
  - R contains additional region and division data per state in the package 'Datasets'
  - This data is from a 1977 report from the chamber of commerce

# External Data

- From census
  - This is a 2016 geography table

# Tail2

-experiment with plotting last 6 rows


```{r tail2, echo=FALSE, size='vsmall', warning=FALSE }
#library(xtable)
#print(xtable(head(AllBeer, 6)))

#library(pander)
#panderOptions('graph.fontsize', .8)
#pander(head(AllBeer, 6))

library(knitr)
#library(kableExtra)
kable(head(AllBeer, 6)) 
#%>%
#kable_styling(font_size = 7)
```


# Simple Plot

```{r plot1, echo=FALSE}
plot(ABV~IBU, data=AllBeer)
abline(lm(ABV~IBU, data=AllBeer))
```


# Fancier Plot

```{r plot2, echo=FALSE, warning=FALSE}
library(ggplot2)

ggplot(dat=AllBeer, aes(x=IBU, y=ABV)) + 
  geom_point(shape=16) + 
  geom_smooth(method=lm) + theme_minimal() 
```

